Jane, John … Leslie? A Historical Method for Algorithmic Gender Detection

### 2014-09-18 article.Rmd
AHR first chart, chart theme

```{r load-libraries, include=FALSE}
library(genderdata)
library(gender)
library(dplyr)
library(ggplot2)
library(mullenMisc)
library(knitr)
library(stringr)
library(scales)

@@

# Load the data frames with the predicted gender, if they exist. Otherwise run
# the analysis.
if(file.exists("data/diss-predicted.rda") &
     file.exists("data/ahr-predicted.rda")) {
  load("data/diss-predicted.rda")
  load("data/ahr-predicted.rda")
} else {
  source("R/load-data.R")
  source("predict-gender.R")
}
```

@@

![Number of dissertations by gender](img/plot-gender.png)

@@

![Proportion of dissertations for each gender](img/dissertations-by-gender-proportion.png)

@@

![Gender of authors of books reviewed in the AHR](img/book-authors.jpg)

>>>>

```{r load-libraries, include=FALSE}
options(stringsAsFactors = FALSE)
library(genderdata)
library(gender)
library(dplyr)
library(ggplot2)
library(mullenMisc)
library(knitr)
library(stringr)
library(scales)
library(tidyr)

@@

# Load the data frames with the predicted gender, if they exist. Otherwise run
# the analysis.
if(file.exists("data/diss-predicted.rda") &
     file.exists("data/ahr-predicted.rda")) {
  load("data/diss-predicted.rda")
  load("data/ahr-predicted.rda")
} else {
  source("R/load-data.R")
  source("predict-gender.R")
}

diss_predicted$gender_author <- diss_predicted$gender_author %>% as.character()
ahr_predicted$gender_author <- ahr_predicted$gender_author %>% as.character()
ahr_predicted$gender_reviewer <- ahr_predicted$gender_reviewer %>% as.character()

diss_predicted[is.na(diss_predicted)] <- "unknown"
ahr_predicted[is.na(ahr_predicted)] <- "unknown"

# Specify article wide chart options
short_scale <- scale_x_continuous(breaks = seq(1950, 2010, 10),
                                  labels = seq(1950, 2010, 10),
                                  limits = c(1950, 2012))

article_theme <- theme(legend.title = element_blank())
```

@@

**Do we even need this chart, or is the chart of proportions sufficient?**

```{r}
diss_predicted %>%
  ggplot(aes(x = year, fill = gender_author)) + 
  geom_bar(binwidth = 1) +
  ggtitle("Dissertations about History by Gender, 1950-2012") +
  xlab(NULL) + ylab(NULL) +
  short_scale + theme(legend.position = "bottom") +
  scale_fill_discrete(name = "",
                      labels = c("Female", "Male")) +
  article_theme
```

@@

```{r}
diss_predicted %>%
  select(year, gender_author) %>%
  filter(year >= 1950) %>%
  group_by(year, gender_author) %>% 
  summarize(n = n()) %>% 
  spread(gender_author, n, fill = 0) %>% 
  mutate(female_proportion = female / (female + male + unknown),
         male_proportion = male / (female + male + unknown),
         unknown_proportion = unknown / (female + male + unknown)) %>% 
  select(female_proportion, male_proportion, unknown_proportion, year) %>%
  gather(gender, proportion, -year) %>% 
  ggplot(aes(x = year, y = proportion, fill = gender)) +
  geom_bar(binwidth = 1, stat = "identity") +
  ggtitle("Proportion of Dissertations about History by Gender, 1950-2012") +
  xlab(NULL) + ylab(NULL) +
  short_scale + theme(legend.position = "bottom") +
  geom_hline(yintercept = 0.5) +
  scale_fill_discrete(labels = c("Female", "Male", "Unknown")) +
  article_theme
```

@@

```{r}
ahr_predicted %>%
  select(year, gender_author) %>%  
  group_by(year, gender_author) %>% 
  summarize(n = n()) %>% 
  spread(gender_author, n, fill = 0) %>%
  mutate(female_proportion = female / (female + male + unknown),
         male_proportion = male / (female + male + unknown),
         unknown_proportion = unknown / (female + male + unknown)) %>%  
  select(female_proportion, male_proportion, unknown_proportion, year) %>%
  gather(gender, proportion, -year) %>% 
  ggplot(aes(x = year, y = proportion, fill = gender)) + 
  geom_bar(binwidth = 1, stat = "identity") +
  ggtitle("Authors of Books Reviewed in the AHR by Gender") +
  xlab(NULL) + ylab(NULL) +
  theme(legend.position = "bottom") +
  scale_fill_discrete(labels = c("Female", "Male")) +
  article_theme
```
